---
title: "Correlation and Confusion Matrix on Iris Data"
output: pdf_document
---

```{r setup, include=FALSE}
if(!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,data.table,reshape)
search()
```

```{r}
library(datasets)
iris.dt <- as.data.table(iris)
```

**Correlation plot using the IRIS data. Which variables have the highest correlation coefficient?**
```{r}
corr <- round(cor(iris.dt[ , c(1:4)]),2)
corr

melted.corr <- melt(corr) 
head(melted.corr)

ggplot(melted.corr, aes(x = X1, y = X2, fill = value)) + 
  scale_fill_gradient(low="wheat", high="orangered") +
  geom_tile() + labs(x = NULL, y = NULL) +
  geom_text(data = melted.corr, aes(x = X1, y = X2, label = value)) +
  ggtitle("Which Variables Are Highly Correlated?")

```

|       Petal.Length and Petal.Width have the highest correlation with 0.96. Petal.Length and Petal.Width also have a high correlation with Sepal.Length having 0.87 and 0.82 correlation coefficients. 


**Calculate average (mean) values of the numeric variables in the data using data.table package. Which variable has the highest mean?**
```{r}

avg <- iris.dt[, sapply(.SD, mean), .SDcols = 1:4]
avg
```
|       Sepal.Length has the highest mean.

**Scatterplot showing the relationship between Sepal.Length and Sepal.Width variable, using ggplot2 package. Color code the points using Species variable.**
```{r}
ggplot(data = iris.dt) + 
  geom_point(mapping = aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  ggtitle("Sepal.Width vs. Sepal.Length ")
```


**Scatterplot showing the relationship between Petal.Length and Petal.Width variables, using ggplot2 package. Color code the points using Species variable.**
```{r}
ggplot(data = iris.dt) + 
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  ggtitle("Petal.Width vs. Petal.Length ")
```
**Combination of variables, creates better separation among records of different Species.**

The plot comparing Petal.Length and Petal.Width create a greater separation among records. There is much less overlap betwwen species versicolor and virginica in this plot as cmpared to the plot of Sepal.Length vs Sepal.Width.



